### 2008-09-20 trunk/htdocs/sections.tex
Added notes on capillary conductance
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@414 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  sections.tex
%
%  Content for Ludwig documentation.
%
%  $Id: sections.tex,v 1.4 2008-08-29 09:32:59 kevin Exp $
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%  (c) 2008 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Capillaries}

The utility \texttt{capillary.c} can be used to generate various
porous media input files corresponding to various simple square
or circular capillaries.

@@

@@

\bibitem{succi}
S. Succi, \textit{The lattice Boltzmann equation and beyond},
2001.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  sections.tex
%
%  Content for Ludwig documentation.
%
%  $Id: sections.tex,v 1.5 2008-09-20 10:39:58 kevin Exp $
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%  (c) 2008 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Example: circular capillaries}

The utility \texttt{capillary.c} can be used to generate various
porous media input files corresponding to various simple circular
or square capillaries.

For an infinite  capillary of circular cross section radius $a$,
the conductance is known analytically\cite{papanastasiou}. The
flow per unit area $J$ is given by
\begin{equation}
J = - \frac{1}{8\eta} \frac{\partial p}{ \partial x} a^2. 
\end{equation}
If the pressure gradient is replaced by a uniform body force
$-\partial p/ \partial x = \rho g$, then one can define a
viscosity-independent conductance $C$ via $ J = C \rho g / \eta$,
i.e., $C = a^2/8$. So, by measuring the rate of change of flow
with applied body force, one can compute an estimate of the conductance
to compare with this result.

If the total momentum is measured from the code for a given force
(in say the $x$-direction), then the flux per unit area is
$J = \sum_i \rho v_x / \sum_i \rho = \sum_i \rho v_x / V$
where the sum is over all fluid lattice sites $i$, with $V$
the fluid volume. The rate of change of flux with applied force
$\Delta J / \Delta\rho g$ can then be combined with the viscosity
to give the conductance.

\subsubsection{Example: rectangular capillaries}

An analtyical expression is also available for the conductance
of a square or rectugular capillary. For an infinite capillary
of square cross section width $w \times h$ [WHICH IS LONGER?],
the volume flux per unit area $J$ is expected to be
\cite{papanastasiou,edo1}
\begin{equation}
J = - \frac{1}{3\eta} \frac{\partial p}{\partial x} (h/2)^2
\left[
1 - 6(h/w) \sum_{k=1}^{\infty} \frac{\tanh(\alpha_k w/h)}{\alpha_k^5} 
\right]
\end{equation}
where $\alpha_k = (2k - 1)\pi/2$. The pressure gradient
$-\partial p / \partial x$ is replaced by the body force $\rho g$ and
one can difine the viscosity-independnet conductance $C$ via
$J = C\rho g / \eta$. The calculation procedes as above.

@@

\bibitem{papanastasiou}
T. paapnastasiou, G. Georgiou, and A. Alexandrou,
\textit{Viscous Fluid Flow},
CRC Press, Boca Raton, Florida, 2000.

@@

\bibitem{succi}
S. Succi, \textit{The lattice Boltzmann equation and beyond},
Oxford University Press, Oxford, 2001.

\bibitem{edo1}
M. Venuroli and E.S. Boek,
Two-dimensional lattice-Boltzmann simulations of single phase
flow in a pseudo two-dimensional micromodel,
\textit{Physica A}, \textbf{362}, 23, 2006.
